---
title: "Homework 3"
author: "Joshua Hummell"
date: "2/17/2021"
output: html_document
---

## Working with Regex  
#### In this assignment, there are four questions that need to be answered!  

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```  
First step is to load my packages:
```{r message=FALSE}
library(dplyr)
library(ggplot2)
library(stringr)
```
  
Now let's begin  
  
  
# #1
#### Using the 173 majors listed in fivethirtyeight.com’s College Majors dataset [https://fivethirtyeight.com/features/the-economic-guide-to-picking-a-college-major/], provide code that identifies the majors that contain either "DATA" or "STATISTICS"
  
Let's load the data!  
```{r}
degrees <- read.csv("https://raw.githubusercontent.com/fivethirtyeight/data/master/college-majors/majors-list.csv",header = TRUE, sep = ",")

str(degrees)
```
  Now we have to look for all the majors that contain "DATA" or "STATISTICS"
```{r}
grep(pattern = 'data|statistics',degrees$Major, value = TRUE, ignore.case = TRUE)
```
And there are only three, how disappointing!

  
    
      



# #2 
#### Write code that transforms the data below:

[1] "bell pepper"  "bilberry"     "blackberry"   "blood orange"

[5] "blueberry"    "cantaloupe"   "chili pepper" "cloudberry"  

[9] "elderberry"   "lime"         "lychee"       "mulberry"    

[13] "olive"        "salal berry"

Into a format like this:

c("bell pepper", "bilberry", "blackberry", "blood orange", "blueberry", "cantaloupe", "chili pepper", "cloudberry", "elderberry", "lime", "lychee", "mulberry", "olive", "salal berry")
  
So, as I understand the question, it wants me to create a character string and then print it out like I am creating a string.   

The only tricky part, which took more time than I am proud to admit to figure out, is that you need a paste within a paste, ie 'Paste Inception'  
```{r}
l <- c("bell pepper", "bilberry", "blackberry", "blood orange", "blueberry", "cantaloupe", "chili pepper", "cloudberry", "elderberry", "lime", "lychee", "mulberry", "olive", "salal berry")


x<- paste('c(', paste('"',l,'"',sep = "", collapse = ','), sep = "",')')


writeLines(x)

```  



# #3 
#### Describe, in words, what these expressions will match:  

"(.)\1\1"  
It will look for the first letter (that does not start on a new line) at the start and see if it repeats twice afterwards  

"(.)(.)\\2\\1"  
I will look at the first two starting letters (that do not start on a new line) and see if something matches the inverse  

"(..)\1"  
It will at the first two letters (that do not start on a new line) and see if something repeats  

"(.).\\1.\\1"  
It will look at the first letter (that does not start on a new line), a character after it, the first character again, a character after it, and then the first character again  


"(.)(.)(.).*\\3\\2\\1"
it will look at the first three letters (that do not start on a new line), see if there is a number after it repeated 0 or more times, then see if there is something that matches the first pattern in the inverse  



# #4
#### Construct regular expressions to match words that:

#### A) Start and end with the same character.  

I was struggling to figure out this code, so I broke it down:

1. in order to get the first character, we need to find out what the first character is so we use '^(.)' 
2. Next, we need to tell the regex that there are other characters in between, so ".*" and then we want to match the end with to the first group, "\\1$"
  + together that is ".*\\1$"
3. ALmost done, but what happens if the string only has 2 characters? We need to add in one more piece in case it's a string like "oo"
  + This one is easy, it's just like the last one, but without the ".*", so "\\1$"
4. When we combine them we will need to use '|' as an if statement
5. Also, to not confuse the regex, I will make sure I separately bracket them, although it is not needed.
  +It would work the same if it were written as "^(.)(.*\\1$|\\1$)"
  


```{r}
y<- c("bob", "toot", "gag", "oo")


str_view(y, "^(.)((.*\\1$)|\\1$)")

```




#### B) Contain a repeated pair of letters (e.g. "church" contains "ch" repeated twice.)  



The hint on this one is that it asks specifically for a letter. I first developed it as if i were looking for any character
1. Find the pair of **letters** '([A-Za-z][A-Za-z])' - Note, the letters specification **needs** to be in parentheses
2. Look through all the text '.*'
3. Find a match '\\1'



```{r}
yy<-c("church", "toto", "yoyo","appropriate")


str_view(yy, "([A-Za-z][A-Za-z]).*\\1")


```







#### C) Contain one letter repeated in at least three places (e.g. "eleven" contains three "e"s.)  


Again, notice the Letters!  

1. Find the first letter '([A-Za-z])'
2. Look through all the text '.*'
3. Find a match '\\1'
4. repeat steps 3 and 4






```{r}
z<-c("eleven", "believe", "tomorrow","individual")


str_view(z, "([A-Za-z]).*\\1.*\\1")


```




