---
title: "Project 3"
author: "Joshua Hummell, Euclides N. Rodriguez, Jie Zou, Evan McLaughlin"
date: "3/19/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Overview
#### We interpreted the project goal as simply to find out what skills are most sought out for Data Scientists. We figured other teams would be more focused on US results (as is expected), but we wanted to compare the industry with our friends in the UK.  

## The Plan  
1. we are going to extract the data from the HTML from the UKs most common job board website, which is [https://www.reed.co.uk/](https://www.reed.co.uk/).   
2. We are going to Tidy the data  
3. Create an analysis about the data  
4. Run  basic NLP on the descriptions

### Load necesary packages  
```{r message=FALSE}
library(rvest)
library(tidyverse)
```



### Step One

1. Pull in the HTML  
- Joshua Hummell

+ The first thing to do is check out the job listing page itself. For these purposes we are going to run only the first page, and check out what we have. I'm not going to show it here because it will take up way to much space, it is a lot of HTML.  


```{r}
reed_data <- read_html("https://www.reed.co.uk/jobs/data-scientist-jobs-in-london?fulltime=True&proximity=1000")
```
  

2. Get the total number of jobs and how many are on this page.  

+ Since we are going to be looking across multiple pages, we need the count of how many jobs we are currently looking at and how many there are in total. This way we can build a loop that will pull from all pages as long as there are jobs.  
```{r}
job <- reed_data %>% 
  html_nodes("div") %>% 
  html_nodes(xpath = '//*[@class="page-counter"]')%>% 
  html_text() %>%
  strsplit(" ")

current_job <- as.numeric(job[[1]][27])
total_job <- as.numeric(job[[1]][29])
paste('On this page there are', current_job, 'jobs out of a total of', total_job, "jobs")
```

3. Get the jobs URLS
- Our plan is to scrape all the jobs pages individually
+ When I tried to scrape this page there were only partial descriptions.    

+ So we are going to create a loop that will pick out all the 'data-ids', which are used to identify the post page in the url.   


First we need to get the first page.  
```{r}
job_url <- reed_data %>% 
  html_nodes("div") %>% 
  html_nodes(xpath = 'a')%>% 
  html_attr('data-id')
job_url <- job_url[!is.na(job_url)]
head(job_url)
```

Now we can get the rest of the pages 
```{r}
# We already got page one, so we want to start it out on page 2
n_page=2

start_time <- Sys.time()

while (current_job < total_job){
    # This will concatenate the url depending on the page
    p = str_c('https://www.reed.co.uk/jobs/data-scientist-jobs-in-london?pageno=',n_page,'&fulltime=True&proximity=450', sep="")
    URL_p = read_html(p)
    
    # This will get the url
    url <- URL_p %>% 
      html_nodes("div") %>% 
      html_nodes(xpath = 'a')%>% 
      html_attr('data-id')
    url <- url[!is.na(url)]
    
    # This appends the data together
    job_url <- append(job_url, url)
    
    # This gets the new job count and changes current job to that number
    job <- URL_p %>% html_nodes("div") %>%  html_nodes(xpath = '//*[@class="page-counter"]')%>% html_text() %>% strsplit(" ")
    current_job <- as.numeric(job[[1]][27])

    # This tells us to go to the next page
    n_page <- n_page + 1
    
}

end_time <- Sys.time()
paste("There are now", current_job, "jobs out of a total of", total_job, "jobs, and it took" ,round(end_time - start_time), "seconds to complete.")
```
  
4. Now that we have all the urls, we are going to crawl each page and scrape the data and add the data from each page to a data frame.  
- Create the data frame  

```{r}
all_jobs <- data.frame(description=character(),
                       position=character(),
                       posted=as.Date(character()),
                       salary=character(), 
                       location=character(),
                       contract=character(),
                       company=character(),
                       company_type=character(),
                       industry=character(),
                       url=character()) 

```

- Now all we have to do is create a loop that will scrape this information off of each page and fill it into the dataframe.  
```{r warning=FALSE}
start_time <- Sys.time()

for (i in unique(job_url)) {
  p = str_c('https://www.reed.co.uk/jobs/data-scientist/',i, sep="")
  URL_p = read_html(p)
  
  
 # Let's get the description
  Desc <- URL_p %>% html_nodes("[itemprop='description']") %>%
            html_text()
  Desc <- str_trim(Desc, side = "left")

  # Let's get the position
  Pos <- URL_p %>% html_node("title") %>%
            html_text()
  
  # Let's get the posted date
  Post <- URL_p %>% html_nodes("[itemprop='datePosted']") %>%
            html_attr('content')
  
  # Let's get the salary
  Sal <- URL_p %>% html_nodes("[data-qa='salaryLbl']") %>%
            html_text()
  Sal <- str_trim(Sal, side = "left")
  
  # Let's get the location
  Loc <- URL_p %>% html_nodes("[data-qa='regionLbl']") %>%
            html_text()
  
  # Let's get the contract
  Cont <- URL_p %>% html_nodes("[data-qa='jobTypeMobileLbl']") %>%
            html_text()
  
  # Let's get the company name
  Comp <- URL_p %>% html_nodes(css ="[itemprop='hiringOrganization']") %>%
            html_nodes(css ="[itemprop='name']") %>%
            html_text() 
  Comp <- str_trim(Comp, side = "left")
  
  # Let's get the company type. Since it is in the Javascript, we need to use regex to extract the value
  Compt <- URL_p %>% str_extract("(jobRecruiterType: )'(\\w+\\s\\w+\\s\\w+|\\w+\\s\\w+|\\w+|\\s)") %>%
      str_extract("(?<=\\')\\D+") 
  
  # Let's get the Industry. Since it is in the Javascript, we need to use regex to extract the value
  Ind <- URL_p %>% str_extract("(jobKnowledgeDomain: )'(\\w+\\s\\w+\\s\\w+|\\w+\\s\\w+|\\w+|\\s)") %>%
      str_extract("(?<=\\')\\D+") 
  
  url <- p
  
  temp <- c(Desc, Pos, Post, Sal, Loc, Cont, Comp, Compt, Ind, url)
  
  all_jobs <- rbind(temp, all_jobs)
}

end_time <- Sys.time()
paste("Your dataframe has been built and it took",round(end_time - start_time), "minutes to complete.")
```

- Now let's rename the columns  
```{r}
colnames(all_jobs) <- c("description", "position","posted","salary","location","contract","company","company_type","industry", "url") 

```
And now, on to tidying!  


------------------------------------------------------------------------------------------------------------------------------------------------

### Step Two - Tidying the Data
- Euclides N. Rodriguez

1. Remove all rows where data is broken from being a 'featured' job
```{r}
clean_jobs <- all_jobs


clean_jobs <- clean_jobs %>% filter(!grepl("Luton|Wimbledon|London|Enfield Town",salary))
```



2. Clean up of position column
```{r}
clean_jobs$position <- str_remove(clean_jobs$position, "- reed.co.uk")

clean_jobs$position <- str_trim(clean_jobs$position)
```

3. Separating the salary column and extracting the salary value
```{r warning=FALSE}
clean_jobs <- clean_jobs %>% separate(salary, into = c("min.salary", "max.salary"), sep = "-")


#Extract_Numeric was a good function to extract the salary value.  Regex was not necessary
clean_jobs <- clean_jobs %>% mutate(Min.Salary.Lbs = readr::parse_number(as.character(min.salary)))

clean_jobs <- clean_jobs %>% mutate(Max.Salary.Lbs = readr::parse_number(as.character(max.salary)))

#Was not able to figure out regex where I can identify the first space only. 
#Used '0_' to seperate text and numbers in the max.salary column
clean_jobs <- clean_jobs %>% separate(max.salary, into = c("X1", "X2"), sep = "0 ")

clean_jobs <- clean_jobs %>% select(-min.salary, -X1)

clean_jobs <- clean_jobs %>% separate(X2, into = c("X3", "X4", "X5"), sep = ", ")
```

4. Create the final data frame for analysis
```{r}
tidy_jobs <- clean_jobs %>% relocate(Min.Salary.Lbs, .after = c(3)) %>% 
  relocate(Max.Salary.Lbs, .after = c(4)) %>% 
  rename( 'Salary.Period' = c(6)) %>%
  select(-'X4', -'X5')
```

And on to the analysis!
------------------------------------------------------------------------------------------------------------------------------------------------

### Step Three - Analyizing the Data
- Jie Zou  


```{r load-packages, message=FALSE}
library(DT)
library('plotly')
library(ggplot2)
```


*This job name also represents data science, so when so many names are at work, which ones appear most often?*
```{r}
df <- tidy_jobs


positions <- df %>% 
  count(position)

ggplotly(ggplot(positions %>% filter(n > 3), aes(x = reorder(position, -n), y = n)) +
  geom_bar(stat = "identity") + 
  coord_flip() +
  labs(x = "number of position",
       y = "name of position",
       title = "Names of position that more than three companies posted "))
```
*What are the average annual salary of these companies and how many of them offer the same average annual salary?*

- The average annual salary of these companies are calculated and stored under variable named avg_annual_salary. The most common average annual salary is 5000, which means that the 88 companies with the same average provide the basic salary and the maximum salary span of 2 * 5000 = 10000
```{r}
# subset of data: [position], [company], [average_annual_salary], [company_type]
avg_annum <- df %>%
  filter(Salary.Period == "per annum" & !is.na(Salary.Period)) %>%
  mutate(avg_annual_salary = (Max.Salary.Lbs - Min.Salary.Lbs)/2)%>%
  select(position, avg_annual_salary, company, company_type) %>%
  arrange(desc(avg_annual_salary))

# number of companies offer the same average annual salary
count_company <- avg_annum %>% 
  count(avg_annual_salary) %>% 
  arrange(desc(n))

ggplotly(
  ggplot(count_company, aes(x = avg_annual_salary, y = n)) + 
  geom_bar(stat = "identity") +
  labs(y = "number of company",
       title = "average annual salary within the companies")) 
```

*The larger average, the larger span. which companies offer the largest and the smallest annual salary span?*

- According to the calculation, both Harnham and McGregor Boyall provide the largest salary span to the position of Data Scientist and software developer. It seems reasonable. However, Hudson Shribman offers the least salary span in chemist/Analytical chemist/Materials scientist.
```{r}
# min and max of average annual salary
a<-avg_annum %>% filter(avg_annual_salary == max(avg_annual_salary) | avg_annual_salary == min(avg_annual_salary))
datatable(a)
# min and max span of these companies
b<- df %>% filter(((Max.Salary.Lbs-Min.Salary.Lbs)/2 == 40000 | (Max.Salary.Lbs-Min.Salary.Lbs)/2 == 500) & Salary.Period == "per annum") %>% select(-c(description, posted, location, contract,industry, Salary.Period))
datatable(b)
```

*What are the major types and corresponding ratio of these companies?*

- The majority type of job posting companies is recruitment consultancy  
```{r}
library(scales)
c<-df %>% 
  count(company_type) %>%
  mutate(per = n/sum(n), 
         label = str_c(percent(per), "(", n, ")")) %>%
  arrange(desc(n))

ggplot(data = c) +
  geom_bar(aes(x = "", y = per, fill = company_type), stat = "identity", width = 1) + 
  coord_polar("y", start = 0) +
  geom_text(aes(x = 1.7, y = cumsum(per)-per/2, label = label)) +
  theme_void()
```




*Among recruitment consultancy, which company or which companies have the most influence?*
The company is considered influential, so its business scope should be broad. In other words, a company like this has more cooperative work than other companies, which means it will post more jobs. As we can see from the plot, the number of jobs posted by Harnham far exceeds that of other companies
```{r}
# business with wider scope
d1 <- df %>% 
  filter(company_type == "Recruitment consultancy") %>%
  count(company) %>% 
  arrange(desc(n))

ggplot(d1 %>% filter(n > 2), aes(x = reorder(company, -n), y = n)) + 
  geom_bar(stat = "identity")+
  coord_flip() +
  labs(y = "number of job posts",
       title = "recruitment consultancy with at least 2 job posts")
```
*How is the spread of locations?*

we've select the companies with at least 4 job posts in one location. Ro discover the relationship between companies and location, the scatter plot is shown below, where we can see that most of companies have job location in London, some companies provide more than one locations such as Harnham and Blue Pelican.
```{r}
# location of these job distributed by companies with at least 4 job posts
d2 <- df %>% 
  group_by(company)%>%
  count(location) %>%
  filter(n > 3)
  
ggplot(d2, aes(x = location, y = company)) +
  geom_point(aes(color = location)) + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```
And, finally, on to NLP!

------------------------------------------------------------------------------------------------------------------------------------------------

### Step Three - Natural Language Processing
- Evan McLaughlin
  

#### Overview
Below, we undertook some NLP processes to distill the job descriptions down to a few key skills. After reading in and cleaning the column, we convert the column to tokens, ran the tokens against our stop words list, and found the most prevalent words in the job descriptions. Single words don't provide much value in terms of analytical insight, so, after enhancing our stop word list, we next determined the most common word pairings, which proved to have much more analytical value. Data Science and Machine Learning were the overwhelming leaders in terms of in-demand skills. Considering Data Science more represents a group of skills as opposed to a single skill, we can safely conclude that in London Data Scientist job postings, Machine Learning represents the most sought-after skill.
Nevertheless, it's useful to learn about other in-demand skills, so we filtered out Data Science and Machine Learning from our dataset in order to better visualize other popular skills. "Computer Science," "Data Analytics," and the ever-important "Communication Skills" topped the list of sought-after characteristics in London-area Data Scientist job descriptions. We've used some helpful graphics to help illustrate our findings below.

Load the packages  
```{r warning=FALSE}
library(RColorBrewer)
library('wordcloud2')
library(openintro)
library(wordcloud)
library(tidytext)
library(magrittr)
library(tinytex)
library(stringr)
library(ggpubr)
library(knitr)
library(fmsb)
library(DBI)
library(NLP)
library(tm)
```


- NLP Analyzing Job Description  
```{r}
# first, I'll read in and clean up the job descriptions
jobs_df <- all_jobs$description

# It's easier to manipulate this data how we want to by converting it to a tibble
jobs_tbl <- tibble(txt = jobs_df)
#jobs_tbl

#next, let's tokenize the text of the description and execute a word count to get an idea of the most prevalent words. We'll also run the result against a stop words list to exclude words that don't add any value to our analysis such as "the", "and", "that", etc.
token <- jobs_tbl %>%
  unnest_tokens(word, 1) %>%
  anti_join(stop_words)

token_count <- token %>%
  count(word) %>%
  arrange(desc(n))

token_count
```
- Looking at the output above, it will definitely be more useful to take a look at the most common word pairs, given many of these words are more descriptive in combination with others


```{r}
token_pairs <- jobs_tbl %>%
  unnest_tokens(pairs, 1, token = "ngrams", n = 2)
token_pairs %>%
  count(pairs) %>%
  arrange(desc(n))
```


- Now, let's run the pairs against the stop_word database by separating the pairs and eliminating cases where either word appears in the stop_word list  

```{r}
pairs_separated <- token_pairs %>%
  separate(pairs, c("word1", "word2"), sep = " ")
pairs_df <- pairs_separated %>%
  filter(!word1 %in% stop_words$word) %>%
  filter(!word2 %in% stop_words$word)
pairs_count <- pairs_df %>% 
  count(word1, word2, sort = TRUE)
head(pairs_count)
```


- Before uniting these columns, let's quickly go through the prominent words and eliminate more terms that don't add much value by augmenting the stop_words list and running another.   
- Some such words are job titles, recruiter names, job locations, salary information, contract lengths, etc.  
- We can add to this list if we happen to see any additional words that aren't helpful to our analysis.   


```{r}
new_stop <- data.frame(word = c("apply", "london", "remote","remotely", "interview", "salary", "contract", "candidate", "scientist", "scientists", "team", "analyst", "engineer", "engineers", "manager", "managers", "senior", "employment", "experienced", "consultant", "junior", "month", "level", "masters", "rosie", "months", "experience", "level", "orientation", "opportunity", "principal", "benefits", "nick", "days", "day", "role", "francesca", "goldman", "luke", "anna", "date", "charlotte", "driven"), lexicon = "custom")
my_stopwords <- rbind(new_stop, stop_words)
pairs_df <- pairs_separated %>%
  filter(!word1 %in% my_stopwords$word) %>%
  filter(!word2 %in% my_stopwords$word)
# Let's now reunite the columns into a single pairs for analysis.
pairs_united <- pairs_df %>%
  unite(term, word1, word2, sep = " ")
df_terms <- pairs_united$term
terms_tbl <- tibble(txt = df_terms)
united_count <- pairs_united %>% 
  count(term, sort = TRUE)
head(united_count)
```




- To facilitate visualization, we can narrow down to the most relevant job skills that employers are looking for by setting a floor on the number of instances and condense our data frame.  
```{r}
a <- 30
data <- united_count
Results<-dplyr::filter(data, data[,2]>a)
colnames(Results)<-c("term", "frequency")
ggplot2::ggplot(Results, aes(x=term, y=frequency, fill=term)) + geom_bar(width = 0.75,  stat = "identity", colour = "black", size = 1) + coord_polar(theta = "x") + xlab("") + ylab("") + ggtitle("Term Frequency (min: 30)") + theme(legend.position = "none") + labs(x = NULL, y = NULL)
plotly::ggplotly(ggplot2::ggplot(Results, aes(x=term, y=frequency, fill=term)) + geom_bar(width = 0.75, stat = "identity", colour = "black", size = 1) + 
xlab("") + ylab("") + ggtitle("Word Frequency (min: 30)") + theme(legend.position = "none") + labs(x = NULL, y = NULL) + theme(plot.subtitle = element_text(vjust = 1), plot.caption = element_text(vjust = 1), axis.text.x = element_text(angle = 90)) + theme(panel.background = element_rect(fill = "honeydew1"), plot.background = element_rect(fill = "antiquewhite")))%>% config(displaylogo = F) %>% config(showLink = F)
```



- Data Science and Machine Learning are obviously the overwhelming results, relative to other skills pairs. Data Science is sort of a catch-all term that we should strive to ignore moving forward.  
- Keeping these two terms in the visualization makes it difficult to analyze the remaining results. So let's insert a maximum records constraint in the above graphics to try to add some nuance to our analysis.

```{r}
a <- 30
b <- 100
data2 <- united_count
Results2<-dplyr::filter(data2, data2[,2]>a, data2[,2]<b )
colnames(Results2)<-c("term", "frequency")
ggplot2::ggplot(Results2, aes(x=term, y=frequency, fill=term)) + geom_bar(width = 0.75,  stat = "identity", colour = "black", size = 1) + coord_polar(theta = "x") + xlab("") + ylab("") + ggtitle("Term Frequency (min: 30, max: 100)") + theme(legend.position = "none") + labs(x = NULL, y = NULL)
plotly::ggplotly(ggplot2::ggplot(Results2, aes(x=term, y=frequency, fill=term)) + geom_bar(width = 0.75, stat = "identity", colour = "black", size = 1) + xlab("") + ylab("") + ggtitle("Word Frequency (min: 30, max: 100)") + theme(legend.position = "none") + labs(x = NULL, y = NULL) + theme(plot.subtitle = element_text(vjust = 1), plot.caption = element_text(vjust = 1), axis.text.x = element_text(angle = 90)) + theme(panel.background = element_rect(fill = "honeydew1"), plot.background = element_rect(fill = "antiquewhite")))%>% config(displaylogo = F) %>% config(showLink = F)
```


- Let's just visualize the remaining (ex. Data Science and Machine Learning) terms once more.  
```{r}
c <- 10
d <- 600
Results3<-dplyr::filter(data2, data2[,2]>c, data2[,2]<d)
wordcloud2(Results3, color = "random-light", backgroundColor = "grey", size = 1.75)
```







