---
title: "Tidying and Tansforming Data"
author: "Joshua Hummell"
date: "03/07/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(tidyverse)
```

#### For this assigmnent we have 3 steps:
1. Create a .CSV file (or optionally, a MySQL database!) that includes all of the information above. You’re encouraged to use a “wide” structure similar to how the information appears above, so that you can practice tidying and transformations as described below.  
2. Read the information from your .CSV file into R, and use tidyr and dplyr as needed to tidy and transform your data.  
3. Perform analysis to compare the arrival delays for the two airlines.  

## Step 1 & Step 2  
#### Since I already created the csv file and published it to Github, we are just going to go ahead and pull it in from Github and take a view. 
```{r}
airlines <- read.csv("https://raw.githubusercontent.com/jhumms/DATA607/main/assignment-5/airlines.csv",header = TRUE, sep = ",")
head(airlines)
```  
  
#### So it has some missing values, the first two column names as well as missing data in the columns and the third row which is empty. Let's fix those complainets.  

1. Blank row - for this we will remove it if it is empty or null  
  
```{r}
airlines <- airlines[!apply(is.na(airlines) | airlines == "", 1, all),]
```
  
2. Rename the columns, and we will make them lowercase as well as make any seperated by a '.' with a '_'  

```{r}
airlines <- airlines %>% rename(carrier = 1, status = 2)
airlines <- airlines %>% rename_all(tolower)
airlines <- rename_with(airlines, ~ tolower(gsub(".", "_", .x, fixed = TRUE)))
```
  
3. Add in any missing data using Zoo's na.locf which adds the first non na above into the na.  

(Unfortunately fill() does not work in this scenario becuase you are working with strings instead of numbers, and thanks to [IRTFM on Stackoverflow for the answer](https://stackoverflow.com/questions/52767323/is-there-a-equivalent-for-the-tidyr-fill-for-strings-in-r))  

```{r}
airlines[airlines==""] <- NA
airlines[1] <- lapply(airlines[1], zoo::na.locf)
```
  
Let's take a look at the data now  
```{r}
head(airlines)
```
  
Much Better!  


#### Now we should be good to go to turn this into long format!  
```{r}
library("reshape2")
melt.airlines <- melt(airlines)
melt.airlines <- melt.airlines %>% rename(destination = 3, amt = 4)
head(melt.airlines)

```
  
## Step 3  
### Let's analyze these two and compare arrival delays for these two airlinesby creating a bar chart to see their on time as well as delayed flights.    



```{r}

ggplot(data=melt.airlines, aes(x=carrier, y=amt, fill=status)) +
  geom_bar(stat="identity", position=position_dodge())
```
  
#### It looks like AM West has a larger number of flights and also a larger number of delays. But, to be fair, let's compare the overall delays with the total number of flights for each carrier.   

```{r}
delays <- melt.airlines %>% select(carrier, status, amt) %>% group_by(carrier, status) %>% summarise(amt = sum(amt))
```
  
#### Then we need to cast it so we can get delayed next to on time  
```{r}
cast.delays <- dcast(delays, carrier~...)
cast.delays <- rename_with(cast.delays, ~ gsub(" ", "_", .x, fixed = TRUE))
```
  
#### Finally we need to get the percent of delayed by the total number of flights and show the results!   

```{r}
cast.delays <- cast.delays %>% mutate(perc_delayed = delayed/(delayed + on_time))


ggplot(data=cast.delays, aes(x=carrier, y=perc_delayed)) +
  geom_bar(stat="identity")
```


#### It looks like Alaska has a higher number flights overall, but how about analyzing the delayed flights by destinations?  

#### Let's begin by casting the data so we can see on time and delayed by destination  

```{r}
cast.airlines <- dcast(melt.airlines, carrier+destination~...)

cast.airlines <- rename_with(cast.airlines, ~ gsub(" ", "_", .x, fixed = TRUE))
```

#### Now let's calculate the percent of time a flight to a destination was delayed  
```{r}
cast.airlines <- cast.airlines %>% mutate(perc_delayed = round(delayed/((on_time+ delayed)), 2))

on_time_airlines<- cast.airlines %>% select(carrier, destination, perc_delayed)
```

#### And let's graph it  

```{r}
ggplot(data=on_time_airlines, aes(x=destination, y=perc_delayed, fill=carrier)) +
  geom_bar(stat="identity", position=position_dodge())
```

#### As the reader can see, AM WEST has a higher percentage of delays for each destination, but from what we saw above, Alaska has a higher percentage of delayed flights.  








