---
title: "Text Mining"
author: "Joshua Hummell"
date: "4/17/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Today's Assignment is to work on chapter two of [Welcome to Text Mining with R](https://www.tidytextmining.com/index.html) and extend the code in two ways:  
1. Work with a different corpus of your choosing, and  
2. Incorporate at least one additional sentiment lexicon (possibly from another R package that you’ve found through research).  

The code can be found in [Chapter Two](https://www.tidytextmining.com/sentiment.html) of the book.  


Citation: 
*    Title: Sentiment analysis with tidy data
*    Author:  Julia Silge and David Robinson
*    Date: 4/6/21
*    Code version: a912425
*    Availability: [https://github.com/dgrtwo/tidy-text-mining/blob/master/02-sentiment-analysis.Rmd](https://github.com/dgrtwo/tidy-text-mining/blob/master/02-sentiment-analysis.Rmd)


#### Step one  
The first thing we are going to need to do is install the sentiment datasets  

```{r message=FALSE}
library(tidytext)
#install.packages('textdata')
library('textdata')
get_sentiments("afinn")
```

```{r message=FALSE}
get_sentiments("bing")
```

```{r message=FALSE}

get_sentiments("nrc")
```


The next step is to pick an author. I like F. Scott Fitzgerald, and as a cynical writer, I want to see if the sentiment analysis picks up on that. 
I will download the books from Gutenberg (the ones out of copyright at least).  

```{r}
library(stringr)
#install.packages('gutenbergr')
library('gutenbergr')

## This shows us the works by F Scott
gutenberg_works(str_detect(author, "Fitzgerald, F. Scott"))

paradise <- gutenberg_download(805)
paradise$book <- 'This Side of Paradise'

flappers <- gutenberg_download(4368)
flappers$book <- 'Flappers and Philosophers'

tales <- gutenberg_download(6695)
tales$book <- 'Tales of the Jazz Age'

beautiful <- gutenberg_download(9830)
beautiful$book <- 'The Beautiful and Damned'

#this one wasn't shown in the dataset, so I had to find it online
gatsby <- gutenberg_download(64317)
gatsby$book <- 'The Great Gatsby'


books <- rbind(
                paradise,
                flappers,
                tales,
                beautiful,
                gatsby
                )

```



Now that we have the books in a dataframe, let's add in line number and chapter number for each word.  
```{r}

library(dplyr)


tidy_books <- books %>%
  group_by(book) %>%
  mutate(
    linenumber = row_number(),
    chapter = cumsum(str_detect(text, 
                                regex("^chapter [\\divxlc]", 
                                      ignore_case = TRUE)))) %>%
  ungroup() %>%
  unnest_tokens(word, text)
```

Let's see how this looks and get all the angry/bitter words that are used in the Great Gatsby  
```{r}
nrc_joy <- get_sentiments("nrc") %>% 
  filter(sentiment == "anger")

tidy_books %>%
  filter(book == "The Great Gatsby") %>%
  inner_join(nrc_joy) %>%
  count(word, sort = TRUE)
```


Now we can assign a sentiment for every 50 since these books are typically smaller than the books of Jane Austen   
```{r}
library(tidyr)

f_scott_sentiment <- tidy_books %>%
  inner_join(get_sentiments("bing")) %>%
  count(book, index = linenumber %/% 50, sentiment) %>%
  pivot_wider(names_from = sentiment, values_from = n, values_fill = 0) %>% 
  mutate(sentiment = positive - negative)
```

Let's graph the results  
```{r}
library(ggplot2)

ggplot(f_scott_sentiment, aes(index, sentiment, fill = book)) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~book, ncol = 2, scales = "free_x")
```
I'm curious, did F. Scott become more cynical over time? 
```{r}
f_scott_sentiment %>% group_by(book) %>% summarise(total_sentiment = sum(sentiment)) %>% arrange(desc(total_sentiment))
```
I would have thought yes, but according to this model, F. Scott actually became less cynical over time. His two early books being the most negative.  





Now, let's look at the Great Gatsby individually  
```{r}
great_gastby <- tidy_books %>% 
  filter(book == "The Great Gatsby")
```


Let's see how all the models compare when looking at the book  
```{r}
afinn <- great_gastby %>% 
  inner_join(get_sentiments("afinn")) %>% 
  group_by(index = linenumber %/% 80) %>% 
  summarise(sentiment = sum(value)) %>% 
  mutate(method = "AFINN")

bing_and_nrc <- bind_rows(
  great_gastby %>% 
    inner_join(get_sentiments("bing")) %>%
    mutate(method = "Bing et al."),
  great_gastby %>% 
    inner_join(get_sentiments("nrc") %>% 
                 filter(sentiment %in% c("positive", 
                                         "negative"))
    ) %>%
    mutate(method = "NRC")) %>%
  count(method, index = linenumber %/% 80, sentiment) %>%
  pivot_wider(names_from = sentiment,
              values_from = n,
              values_fill = 0) %>% 
  mutate(sentiment = positive - negative)
```

And let's see how that looks mapped out 
```{r}
bind_rows(afinn, 
          bing_and_nrc) %>%
  ggplot(aes(index, sentiment, fill = method)) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~method, ncol = 1, scales = "free_y")
```





And let's add a wordcloud to see the most popular words in the book  
```{r}
library(wordcloud)

great_gastby %>%
  anti_join(stop_words) %>%
  count(word) %>%
  with(wordcloud(word, n, max.words = 100))
```
```{r}
library(reshape2)

great_gastby %>%
  inner_join(get_sentiments("bing")) %>%
  count(word, sentiment, sort = TRUE) %>%
  acast(word ~ sentiment, value.var = "n", fill = 0) %>%
  comparison.cloud(colors = c("gray20", "gray80"),
                   max.words = 100)
```




















